---
title: "HW1"
author: "Federico Travascio"
date: "2024-12-02"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

**COLLABORATIONS**

I used artificial intelligence as support for the R-codes. In addition,
during the course of the homework I worked with Flavia Scarfò, Ramona
Fantone, Marco Cavallo and Fabio Tripodi.

**EXERCISE 1**

**A.** The first-order autoregressive process (AR(1)) is defined by the
relationship:

$$
X_t = \theta X_{t-1} + \epsilon_t
$$

The AR(1) model can exhibit either stationary or non-stationary
behavior. For the process to be stationary, the following condition must
be met: $|\theta| < 1$.

A process is weakly stationary (or covariance stationary) if it
satisfies three key conditions:

1.  The mean $E[X_t]$ is constant over time. That is, $E[X_t] = \mu$ for
    all $t$.

2.  The variance $\text{Var}(X_t)$ is constant over time. In other
    words, $\text{Var}(X_t) = \sigma_X^2$ for all $t$.

3.  The autocovariance $\text{Cov}(X_t, X_{t+h})$ depends only on the
    lag $h$, not on the time $t$. Specifically, the autocovariance
    function is $\gamma(h) = \text{Cov}(X_t, X_{t+h})$, which depends
    only on the lag $h$.

We will now demonstrate how these conditions apply to our AR(1) process.

**The mean of** $X_t$ is given by:

$$
E[X_t] = E[\theta X_{t-1} + \epsilon_t]
$$

This step uses the definition of the process $X_t$, which is defined as
$\theta$ times the previous value of the process $X_{t-1}$, plus a
random error term $\epsilon_t$. The expected value (or mean) of $X_t$ is
thus equal to the expected value of this expression.

Using the Law of Total Expectation, we get:

$$
E[X_t] = \theta E[X_{t-1}] + E[\epsilon_t]
$$

Here we use the law of total expectation, which states that the expected
value of a sum of random variables is equal to the sum of their expected
values. Also, since $\theta$ is a constant, it can be pulled out of the
expectation operator.

$$
E[X_t] = \theta E[X_{t-1}]
$$

Now, we consider the fact that $\epsilon_t$ is white Gaussian noise with
zero mean ($E[\epsilon_t] = 0$). Therefore, the term $E[\epsilon_t]$
vanishes from the equation.

We assume that the process is stationary, meaning its mean is constant
over time ($E[X_t] = E[X_{t-1}] = \mu$). Substituting $\mu$ for $E[X_t]$
and $E[X_{t-1}]$, we obtain this equation:

$$
\mu = \theta \mu
$$

Therefore, the mean of the process is constant and equal to 0. If $\mu$
were different from zero, then $\theta \mu$ would differ from $\mu$, and
the equation would not hold. Therefore, the mean of the process is
constant and equal to 0. So, the derivation shows that if a process is
defined as a multiple of its previous value plus a white Gaussian noise
term with zero mean, and the process is stationary, then the mean of the
process must be zero.

**The variance of** $X_t$ is:

$$
Var(X_t) = E[X_t^2] - (E[X_t])^2
$$

This is the starting point of the calculation. The variance of a random
variable is defined as the difference between the expected value of the
square of the variable and the square of its expected value.

Since $E[X_t] = 0$, this simplifies to:

$$
Var(X_t) = E[X_t^2]
$$

We substitute $X_t$ with its definition, which is
$\theta X_{t-1} + \epsilon_t$, to expand the square and simplify the
expression.

$$
E[X_t^2] = E[(\theta X_{t-1} + \epsilon_t)^2]
$$

We expand the square of the expression
$(\theta X_{t-1} + \epsilon_t)^2$, obtaining the three terms on the
right-hand side of the equation.

$$
E[X_t^2] = \theta^2 E[X_{t-1}^2] + 2\theta E[X_{t-1} \epsilon_t] + E[\epsilon_t^2]
$$

Since $\epsilon_t$ and $X_{t-1}$ are independent, and
$E[\epsilon_t] = 0$, the term $2\theta E[X_{t-1} \epsilon_t]$ simplifies
to 0.

$$
E[X_t^2] = \theta^2 E[X_{t-1}^2] + E[\epsilon_t^2]
$$

Remembering that $\epsilon_t = Z + \sigma W_t$ , using the Law of Total
Variance.

$$
Var(\epsilon_t) = Var(Z) + \sigma^2
$$

We calculate the variance of $Z$, which is a random variable uniformly
distributed between -5 and 5. The formula for the variance of a uniform
distribution is $\frac{(b - a)^2}{12}$, where $a$ and $b$ are the lower
and upper limits of the interval.

$$
Var(Z) = \frac{(5 - (-5))^2}{12} = \frac{100}{12} = \frac{25}{3}
$$

We substitute the variance of $Z$ calculated in the previous step into
the total variance formula for $\epsilon_t$.

$$
Var(\epsilon_t) = \frac{25}{3} + \sigma^2
$$

Assuming the stationarity of the process, i.e.,
$\text{Var}(X_t) = \text{Var}(X_{t-1}) = \sigma_X^2$, we can substitute
$E[X_t^2]$ and $E[X_{t-1}^2]$ with $\sigma_X^2$ in the equation from
Step 5. Then, we substitute $\text{Var}(\epsilon_t)$ with the result
obtained in Step 8 and solve for $\sigma_X^2$. This provides us with the
formula for the variance of the process $X_t$, which is constant over
time. The condition $|\theta| < 1$ is necessary for the convergence of
the geometric series that forms during the solution for $\sigma_X^2$.

$$
\sigma_X^2 = \theta^2 \sigma_X^2 + \frac{25}{3} + \sigma^2=
$$

$$
 \frac{25}{3} + \frac{\sigma^2}{1 - \theta^2}, \, |\theta| < 1
$$

Assuming the stationarity of the process, i.e.,
$\text{Var}(X_t) = \text{Var}(X_{t-1}) = \sigma_X^2$, we can substitute
$E[X_t^2]$ and $E[X_{t-1}^2]$ with $\sigma_X^2$ in the equation from
Step 5. Then, we substitute $\text{Var}(\epsilon_t)$ with the result
obtained in Step 8 and solve for $\sigma_X^2$. This provides us with the
formula for the variance of the process $X_t$, which is constant over
time. The condition $|\theta| < 1$ is necessary for the convergence of
the geometric series that forms during the solution for $\sigma_X^2$.

**The autocovariance** is:

$\text{Cov}(X_t, X_{t+h}) = \text{Cov}(X_t, \theta^h X_t + \sum_{i=0}^{h-1} \theta^i \epsilon_{t+h-i})$

This is the starting point of the calculation. The covariance between
two random variables measures their tendency to vary together. In this
case, we are trying to calculate the covariance between the process at
time $t$ ($X_t$) and the process at time $t+h$ ($X_{t+h}$). We use the
definition of $X_{t+h}$ provided in the source to expand the expression.

$\text{Cov}(X_t, X_{t+h}) = \text{Cov}(X_t, \theta^h X_t)$

This step takes advantage of the fact that $X_t$ and $\epsilon_{t+h-i}$
are independent for each $i$. The covariance between two independent
variables is always zero. Therefore, all the terms in the summation that
involve $\epsilon_{t+h-i}$ cancel out, leaving only the covariance
between $X_t$ and $\theta^h X_t$.

$\text{Cov}(X_t, X_{t+h}) = \theta^h \text{Cov}(X_t, X_t)$

Now, we apply the property of covariance that states that covariance is
a bilinear operator. This means we can factor out the constant
$\theta^h$ from the covariance.

$\text{Cov}(X_t, X_{t+h}) = \theta^h \text{Var}(X_t)$

The covariance of a random variable with itself is equal to its
variance. Therefore, $\text{Cov}(X_t, X_t)$ is simply the variance of
$X_t$, denoted as $\text{Var}(X_t)$.

$\text{Cov}(X_t, X_{t+h}) = \theta^h \left( \sigma^2 \frac{1}{1 - \theta^2} + \frac{1}{(1-\theta)^2} \cdot \frac{100}{12} \right)$

We substitute $\text{Var}(X_t)$ with its formula, which was previously
derived as
$\sigma^2 \frac{1}{1 - \theta^2} + \frac{1}{(1-\theta)^2} \cdot \frac{100}{12}$.
This gives us the final expression for the covariance between $X_t$ and
$X_{t+h}$.

Since $|\theta| < 1$ (the stationarity condition),
$\text{Cov}(X_t, X_{t+h})$ is finite and tends to 0 as $h \to \infty$,
because $\theta^h \to 0$ when $|\theta| < 1$.

This step analyzes the behavior of the covariance as the time distance
between the two points of the process ($h$) increases. The stationarity
condition, $|\theta| < 1$, implies that $\theta^h$ approaches zero as
$h$ tends to infinity. Therefore, the covariance between $X_t$ and
$X_{t+h}$ decreases to zero as the time distance increases.

The derivation demonstrates that the covariance between two points of
the process $X_t$ separated by a time interval $h$ is finite and
decreases to zero as $h$ tends to infinity. This result confirms the
weak stationarity of the process, meaning that its mean and variance are
constant over time, and its autocovariance depends only on the time
distance between the two points considered.

**B. Ergodicity:**

This is the definition of the autocorrelation function (ACF). The ACF
measures the linear correlation between two points of the stochastic
process separated by a time interval $( h )$. It is defined as the ratio
between the covariance of $( X_t )$ and $( X_{t+h} )$, and the variance
of $( X_t )$.

$( \rho_h = \frac{\text{Cov}(X_t, X_{t+h})}{\text{Var}(X_t)} )$

In this step, we substitute the numerator with the result obtained
earlier for $\text{Cov}(X_t, X_{t+h})$, which is
$\theta^h \text{Var}(X_t)$.

$\rho_h = \frac{\theta^h \text{Var}(X_t)}{\text{Var}(X_t)}$

Simplifying the fraction, we obtain the final expression for the
autocorrelation function: $\rho_h = \theta^h$. This expression shows
that the autocorrelation depends only on $\theta$ and the time interval
$h$.

$\rho_h = \theta^h$

Since $|\theta| < 1$, $\rho_h$ decays exponentially to 0 as $h$
increases.

The derivation shows that the autocorrelation function of the process is
$\rho_h = \theta^h$, and it decays exponentially to zero as $h$
increases. This behavior indicates that the process has "short memory,"
meaning that the influence of the past on the present quickly diminishes
as the time distance increases.

The conclusion asserts that the process is not ergodic because the
autocorrelation function depends on $Z$, which varies between different
realizations of the process. This conclusion is based on the assumption
that $Z$ is different for each realization of the process.

**C.**

```{r}
set.seed(23072001)

# Parameters
sigma2 <- 1  # Variance of the white noise
theta <- 0.9 # Autoregressive coefficient
T <- 100     # Length of the time series

# Generate Z from Uniform[-5, 5]
Z <- runif(1, min = -5, max = 5)

# Generate the white noise as Gaussian noise with mean Z and variance sigma2
epsilon_t <- rnorm(T, mean = Z, sd = sqrt(sigma2))

# Initialize the time series X_t
X_t <- numeric(T)  # Empty vector to store the time series
X_t[1] <- epsilon_t[1]

# Simulate the AR(1) process
for (t in 2:T) {
  X_t[t] <- theta * X_t[t - 1] + epsilon_t[t]
}

# Convert X_t to a time series object
X_t_ts <- ts(X_t)

# Plot the time series
plot(X_t_ts, main = "Simulated AR(1) Process", ylab = "Xt", xlab = "Time", col = "purple")
```

The plot above shows the evolution of a simulated AR(1) process over
time. Below are some key observations regarding its stationarity:

-   The series exhibits fluctuations that appear to stabilize around a
    certain level over time, suggesting that it may oscillate around a
    steady-state mean.
-   In the early portion of the series, there is a noticeable upward
    trend, followed by fluctuations with smaller amplitude in the later
    periods.
-   Despite these short-term trends, there are no strong signs of
    explosive behavior, which would be expected if the process were not
    stationary. However, with only 100 observations, the series does not
    exhibit perfect stationarity to the naked eye.

And also:

-   The theoretical condition for stationarity in an AR(1) process is
    that $|\theta| < 1$.
-   Assuming this series was generated with $\theta = 0.9$, it satisfies
    the condition $|\theta| < 1$, ensuring theoretical stationarity.
-   However, due to the relatively short time series, the properties of
    stationarity may not yet be fully apparent, as the process needs
    more time to "settle" around its long-term mean.

So, in conclusion:

-   Visually, the series does not appear entirely stationary due to the
    sample size and observed trends.
-   Increasing the number of observations would likely reveal clearer
    stationarity, with the process oscillating more evidently around the
    theoretical mean and variance stabilizing.

```{r}
# Compute the empirical mean
empirical_mean <- mean(X_t)
cat("Empirical Mean:", round(empirical_mean,2) , "\n")
```

```{r}
# Compute the empirical variance
empirical_variance <- var(X_t)
cat("Empirical Variance:", round(empirical_variance,2) , "\n")
```

The results, with a mean of $35.96$ and a variance of $50.23$, deviate
significantly from the theoretical values. The theoretical mean of the
process is $0$, but the empirical value can deviate due to the influence
of the specific random value $Z$ used in the realization of the process.
This could explain the significant difference between my mean ($35.96$)
and the theoretical mean ($0$).

Knowing that, the variance $\text{Var}(X_t)$ is given by:\
$$
\text{Var}(X_t) = \frac{\sigma^2}{1 - \theta^2} + \frac{1}{(1 - \theta)^2} \cdot \frac{100}{12},
$$ where $\sigma^2 = 1$ and $\theta = 0.9$.

Substituting these values, we obtain:\
$$
\text{Var}(X_t) = \frac{1}{1 - 0.9^2} + \frac{1}{(1 - 0.9)^2} \cdot \frac{100}{12}.
$$

The theoretical variance can be computed explicitly as:\
$$
\text{Var}(X_t) \approx 838.6,
$$which is significantly larger than the empirical variance $50.23$.
This discrepancy suggests that the simulation generated a series with
less dispersion than predicted by theory.

```{r}
# Number of simulations
num_simulations <- 5  

# Matrix to store multiple time series
time_series_matrix <- matrix(0, nrow = T, ncol = num_simulations)

# Simulate the process for different Z values
for (sim in 1:num_simulations) {
  Z <- runif(1, min = -5, max = 5)  # Draw a new Z for each simulation
  epsilon_t <- rnorm(T, mean = Z, sd = sqrt(sigma2))  # Generate epsilon_t
  X_t <- numeric(T)  # Initialize the time series
  X_t[1] <- epsilon_t[1]
  
  # Generate the AR(1) process
  for (t in 2:T) {
    X_t[t] <- theta * X_t[t - 1] + epsilon_t[t]
  }
  
  # Store the result in the matrix
  time_series_matrix[, sim] <- X_t
}

# Plot all time series on a single plot
matplot(time_series_matrix, type = "l", lty = 1, col = rainbow(num_simulations),
        xlab = "Time", ylab = "Xt", main = "AR(1) Processes with Different Z Values")
```

The results of the plot show that the AR(1) process converges to unique
equilibrium states that depend on the observed value of $Z$. Each
realization of the process, characterized by a different value of $Z$
drawn from the uniform distribution $U[-5,5]$, converges to a distinct
horizontal band in the graph. This behavior is due to the fact that $Z$
determines the mean of the error term $\epsilon_t$, influencing the
equilibrium level of each realization.

However, the dependence on $Z$ implies that the AR(1) process is not
ergodic. Ergodicity requires that the process's statistics, such as the
mean and variance, can be reliably estimated from a single sufficiently
long realization. In the case of the AR(1) process under consideration,
the different values of $Z$ lead to distinct equilibrium states, making
it impossible to obtain accurate estimates of the process's statistics
from a single realization.

```{r}
acf(X_t, lag.max = 50, main = "Autocorrelation Function (ACF)")
```

The graphs show that the process converges to distinct equilibrium
states determined by the value of the random parameter $Z$. Each
realization of the process, characterized by a different value of $Z$,
stabilizes around a specific mean level. This behavior is expected in a
stationary AR(1) process, where past values influence the present, but
the effect gradually diminishes over time.

The observed empirical mean ($35.96$) deviates significantly from the
theoretical mean ($0$). This discrepancy can be attributed to the
influence of the specific random value $Z$ used in the realization of
the process.

The theoretical variance is approximately $838.6$, which is considerably
higher than the empirical variance obtained from the simulation. This
difference may be due to the limited sample size ($T = 100$) or the
specific sequence of random numbers used in the simulation.

A larger sample size would tend to reduce variability in the estimates
and produce results closer to the theoretical values.

The autocorrelation function (ACF) of the AR(1) process decays
exponentially at a rate determined by the parameter $\theta$. A value of
$\theta$ close to $1$ indicates longer memory, while a value of $\theta$
close to $0$ indicates shorter memory. The shape of the ACF provides
information on the persistence of fluctuations in the process.

The sample autocorrelation coefficient $\hat{\rho}_h$ is an estimator of
the theoretical autocorrelation $\rho_h$ for lag $h$. It is computed as:
$$
\hat{\rho}h = \frac{\sum{t=1}^{T-h} (X_t - \bar{X})(X_{t+h} - \bar{X})}{\sum_{t=1}^T (X_t - \bar{X})^2},
$$ where $\bar{X}$ is the sample mean of the time series and $T$ is the
length of the series.

*Observations*

-   *Accuracy of* $\hat{\rho}_h$:
    -   For smaller lags ($h$), the estimator $\hat{\rho}_h$ closely
        approximates the theoretical autocorrelation
        $\rho_h = \theta^h$. This can be observed in the ACF plot, where
        the sample autocorrelation follows the expected exponential
        decay pattern.
    -   As $h$ increases, $\hat{\rho}h$ becomes less reliable due to the
        reduced number of overlapping data points $(X_t, X{t+h})$ and
        the increased influence of noise.
-   *Convergence and Variability*:
    -   The variability of $\hat{\rho}_h$ increases with $h$, as seen
        from the confidence intervals in the ACF plot. This is expected
        since the number of observations contributing to $\hat{\rho}_h$
        decreases with larger $h$, making the estimates noisier.
-   *Relation to* $Z$:
    -   In the specific context of this AR(1) process, the dependence of
        the equilibrium state on the random variable $Z$ introduces
        additional variability into $\hat{\rho}_h$ across different
        realizations of the process. This highlights the non-ergodic
        nature of the process, as the autocorrelation estimates may
        differ between realizations depending on the value of $Z$.

The estimator $\hat{\rho}_h$ is a consistent and unbiased estimator for
$\rho_h$ for stationary processes when $T \to \infty$. In this case, it
effectively captures the exponential decay of autocorrelation for the
AR(1) process with $\theta = 0.9$. However, for finite sample sizes and
larger lags, the estimates become less precise due to limited data and
noise.

**EXERCISE 2**

```{r}
S <- c(

1, 0, 2, 2, 1, 1, 1, 1, 0, 0, 2, 2, 0, 1, 0, 1, 1, 2, 1, 2, 11, 1, 1, 2, 1, 0, 1, 1, 0, 1, 1, 0, 2, 0, 2, 1, 1, 0, 2, 1, 1, 1, 0, 0, 2, 1, 0, 2, 0, 1, 2, 0, 1, 0, 1, 1, 2, 1, 0, 2, 1, 2, 2, 0, 1, 1, 2, 1, 0, 0, 0, 2, 0, 1, 1, 1, 2, 1, 2, 0, 1, 2, 2, 2, 1, 1, 1, 0, 2, 0, 2, 0, 2, 1, 2, 1, 2, 1, 1, 1)
```

**A.** Since tossing two independent coins results in three possible
outcomes, it is logical to define $St$ as the sum of the outcomes from
each coin toss:

-   $St=2$ if both coins show heads (HH)

-   $St=1$ if one coin shows heads and the other tails (HT or TH)

-   $St=0$ if both coins show tails (TT).

These outcomes form a sequence of independent and identically
distributed random variables following a Binomial distribution
$St ​∼ Binomial (n=2, p=0.5)$. we define $Xt$​ as a linear combination of
the past two values $S_{t−1}$ and $S_{t−2}$​, along with the current
value $St$​:

$$
X_t = \alpha_1 S_{t-1} + \alpha_2 S_{t-2} + S_t
$$

Thus, the process is classified as an $MA(2)$ because it incorporates
the current shock $St$​, as well as the two most recent shocks $S_{t−1}$,
$S_{t−2}$, the coefficients α*1 and α*2​ acting as the respective
weights.

The previous process definition derives from the general MA(2) process:
$$
X_t = \theta_0 + \theta_1 \epsilon_{t-1} + \theta_2 \epsilon_{t-2} + \cdots + \theta_q \epsilon_{t-q} + \epsilon_t = 
$$$$ X_t = \theta_0 + \sum_{j=1}^{q} \theta_j \epsilon_{t-j} + \epsilon_t
$$

where:

-   $θ_0$ ​: Constant term representing the mean of the process.

-   $θ_j$​: Coefficients that determine the influence of past shocks
    $ϵt−j​$ on $Xt$​.

-   $ϵt−j$: Past random shocks occurring at time $t−j$, assumed to be
    independent with zero mean and constant variance.

-   $ϵt$​: The current random shock (white noise) at time $t$.

Since there are 3 possible outcomes for each toss, the distribution that
describes the White Noise is a discrete distribution.

**B.** Now, let us calculate that the mean and variance are constant
over time and that the autocorrelation depends on the lag $h$ and not on
the lag $t$. This is a sufficient assumption to define the stationarity
of the process $Xt$:

-   **MEAN**

To compute the expected value of $Xt$, we apply the linearity of
expectation:

$E[X_t] = E[\alpha_1 S_{t-1} + \alpha_2 S_{t-2} + S_t]$

Using the linearity property of expectation:

$$E[X_t] = \alpha_1 E[S_{t-1}] + \alpha_2 E[S_{t-2}] + E[S_t]$$

Considering that:

$E[S_t] = n \cdot p = 2 \cdot 0.5 = 1$

for each $t$, since each coin toss has an equal chance of being heads or
tails, $St$​ follows a Binomial distribution with parameters $n=2$ and
$p=0.5$ we get:

$E[X_t] = \alpha_1 \cdot 1 + \alpha_2 \cdot 1 + 1$

Which can be simplified in this way:

$E[X_t] = \alpha_1 + \alpha_2 + 1$

Therefore, as a condition of stationarity, $E[X_t]$ is constant over
time, depending only on $$\alpha_1$$ and$$\alpha_2$$ and not on $t$.

-   **VARIANCE**

To demonstrate that the variance of $Xt$ is constant over time, we
proceed through the following steps:

1.  The process $Xt$ is given by:

$Xt=α1St−1+α2St−2+StX_t$

\
We start by recalling the definition of $Xt$, which is a linear
combination of the random variables $S_{t−1}$, $S_{t−2}$, and $St$. This
setup suggests that the variance of $Xt$ will depend on the individual
variances of these terms, scaled by their respective coefficients $α_1$,
$α_2$, and 1.

2.  To calculate the variance of $Xt$, we use the formula for the
    variance of a linear combination of random variables. Specifically,
    for any random variables $X$ and $Y$, and constants $a$ and $b$, the
    variance of $aX+bY$ is given by:

$Var(aX+bY)=a2Var(X)+b2Var(Y)\text{Var}(aX + bY)$

Since $St$ is independent across time (because it is white noise), the
variances of the components can be added. Therefore, the variance of
$Xt$ is:

$Var(X_t)=Var(α_1S_{t−1}+α_2S_{t−2}+St)\text{Var}(X_t)$

Using the property of independence and the scaling effect of the
coefficients:

$Var(Xt)=α_1^2Var(S_{t−1})+α_2^2Var(S_{t−2})+Var(St)\text{Var}(X_t)$

\
Considering that $S_{t−1}$, $S_{t−2}$, and $St$ are independent (they
represent separate coin tosses), we can add their individual variances.
The coefficients $α1$ and $α2$ scale the variances of $S_{t−1}$ and
$S_{t−2}$, respectively, while $St$ contributes its variance as is.

3.  Given that St follows a Binomial distribution with parameters $n=2$
    and $p=0.5$, we can calculate its variance as follow:

$Var(St)=n⋅p⋅(1−p)=2⋅0.5⋅0.5=0.5\text{Var}(S_t)$

Thus, for all values of t, the variance of $St$ is constant and equals
0.5:

$Var(S_{t−1})=Var(S_{t−2})=Var(St)=0.5$

The variance of $St$ is constant because it is based on a Binomial
distribution with fixed parameters. This is crucial for the next step
because we can now substitute this constant variance into the variance
formula for $Xt$, ensuring that the variance of $Xt$ is determined by
the fixed coefficients $α1$ and $α2$, and the time-invariant variance of
$St$.

Substituting the value of $Var(St)=0.5$ into the formula for $Var(Xt)$
we get:

$Var(Xt)=α_{1/2}⋅0.5+α_{2/2}⋅0.5+0.5$

$Var(Xt)=0.5(α_{1/2}+α_{2/2}+1)$

At this point, we substitute the constant value of $Var(St)=0.5$ into
our expression for $Var(Xt)$, simplifying the formula to show how the
variance of $Xt$ depends on the fixed coefficients $α_1$ and $α_2$, as
well as the constant variance of $St$.

The variance of $Xt$ is constant over time because it depends only on
the fixed values of $α_1$ and $α_2$, as well as the constant variance of
$St$. Since the variance of $St$ is independent of time, the variance of
$Xt$ does not change over time. Therefore, $Var(Xt)$ satisfies the
condition for weak stationarity, which requires that the variance is
time-invariant.

\
This final step concludes that the variance of $Xt$ is time-invariant,
proving that the process satisfies the weak stationarity condition for
variance. The key factor is that the variance depends only on fixed
parameters and a time-invariant white noise process, which guarantees
that $Xt$ has constant variance.

-   **AUTOCORRELATION FUNCTION**

To prove the stationarity of the process:

**At lag h=1**, the covariance between $Xt$ and $X_{t−1}$ is given by:

Considering that $X_t$ and $X_{t-1}$ share the term $S_{t-1}$, there is
a correlation between them due to the influence of the white noise at
time $t-1$. This correlation arises because the process $S_t$ is
independent at different times, but within the same lag, overlapping
terms contribute to the covariance.

$Cov(X_t,X_{t−1})=Cov(α1S_{t−1}+α1α2S_{t−2})$

Substituting $Var(St)=0.5$, we find:

$Cov(Xt,X_{t−1})=α_1⋅Var(S_{t−1})+α_1α_2⋅Var(S_{t−2})$=

=$α_1⋅0.5+α_1α_2⋅0.5$\

**At lag h=2**, the covariance between $Xt$ and $X_{t−2}$ is given by:

$Cov(Xt,X_{t−2})=Cov(α1S_{t−1}+α_2S_{t−2}+St,S_{t−2}+α_1S_{t−3}+α_2S_{t−4})$=

=$Cov(α_2S_{t−2})$= $=α_2Var(S_{t−2})=α_2⋅0.5$

At lag 2, the covariance involves only the term $α_2S_{t−2}$, since only
$S_{t−2}$ is shared between $Xt$ and $X_{t−2}$. The other terms do not
contribute due to independence. We again use the constant variance of
$St$, which is 0.5, to calculate the covariance.

**For h\>2**, the covariance is zero:

$Cov(Xt,X_{t−h})=0$

In this case, no terms overlap between $Xt$ and $Xt−h$ because the time
indices are too far apart. Since $St$ is white noise and independent at
different time points, the covariance for these higher lags is zero.
Consequently, autocovariance will be zero.

Finally we can find the **autocorrelation function** $ρ(h)$, we use the
autocovariance $γ(h)=Cov(Xt,Xt−h)$, which we derived earlier. The ACF is
defined as:

$$
\rho(h) = \frac{\gamma(h)}{\gamma(0)}
$$

where $\gamma(0)$ is the variance of $X_t$, and $\gamma(h)$ is the
covariance between $X_t$ and $X_{t-h}$.

The covariance at **lag h = 1** is:

$$
\gamma(1) = \text{Cov}(X_t, X_{t-1}) = \text{Cov}(\alpha_1 S_{t-1} + \alpha_2 S_{t-2} + S_t, \alpha_1 S_{t-2} + \alpha_2 S_{t-3} + S_{t-1})
$$

Only overlapping terms contribute due to the independence of $S_t$:

$$
\gamma(1) = \alpha_1 \cdot \text{Var}(S_{t-1}) + \alpha_1 \alpha_2 \cdot \text{Var}(S_{t-2})
$$

Since $\text{Var}(S_t) = 0.5$:

$$
\gamma(1) = \alpha_1 \cdot 0.5 + \alpha_1 \alpha_2 \cdot 0.5=
$$

$$ = 0.5 \alpha_1 (1 + \alpha_2)$$

Thus, the autocorrelation at lag 1 becomes:

$$
\rho(1) = \frac{\gamma(1)}{\gamma(0)} = \frac{0.5 \alpha_1 (1 + \alpha_2)}{0.5 (1 + \alpha_1^2 + \alpha_2^2)}=
$$

$$
= \frac{\alpha_1 (1 + \alpha_2)}{1 + \alpha_1^2 + \alpha_2^2}
$$

For lag **h = 2**, the covariance is:

$$
\gamma(2) = \text{Cov}(X_t, X_{t-2}) = \text{Cov}(\alpha_1 S_{t-1} + \alpha_2 S_{t-2} + S_t, S_{t-2} + \alpha_1 S_{t-3} + \alpha_2 S_{t-4})
$$

Only the term $S_{t-2}$ overlaps:

$$
\gamma(2) = \alpha_2 \cdot \text{Var}(S_{t-2}) = \alpha_2 \cdot 0.5
$$

Thus, the autocorrelation at lag 2 becomes:

$$
\rho(2) = \frac{\gamma(2)}{\gamma(0)} = \frac{0.5 \alpha_2}{0.5 (1 + \alpha_1^2 + \alpha_2^2)}=
$$

$$
= \frac{\alpha_2}{1 + \alpha_1^2 + \alpha_2^2}
$$

In order to calculate lags greater than two, we take **lag = 3** as a
reference$$
\gamma(h) = \text{Cov}(X_t, X_{t-h}) = 0 \quad \text{for} \quad h > 2
$$

Thus, the autocorrelation is:

$$
\rho(h) = \frac{\gamma(h)}{\gamma(0)} = \frac{0}{\gamma(0)} = 0
$$

\
Since $X_t$ and $X_{t-h}$ share no terms in common for $h > 2$.

In the end, we demonstrated that the mean and the variance of $X_t$ are
constant and the autocorrelation function $\rho(h)$ shows that the
correlation between $X_t$ and $X_{t-h}$ depends only on the lag $h$,
confirming that $X_t$ is **weakly stationary**.

### 

**C.** Since the process in question is an MA(2), we know that the
theoretical PACF will be zero for $h > 2$. Therefore, we only need to
calculate $\Phi_{11}$ and $\Phi_{22}$.

-   $\Phi_{11}$ is simply the correlation between $X_t$ and $X_{t+1}$,
    which is equal to $\rho_1$:

For $h = 1$, the PACF is:

$$
\phi_{11} = \rho_1.
$$

But we already know the value of $\rho_1$:

$$
\rho_1 = \frac{0.5 \alpha_1 + 0.5 \alpha_1 \alpha_2}{0.5(1 + \alpha_1^2 + \alpha_2^2)}.
$$

So, canceling out 0.5, we will have:

$$
\phi_{11} = \frac{\alpha_1 (1 + \alpha_2)}{1 + \alpha_1^2 + \alpha_2^2}.
$$

#### PACF for $h = 2$

The PACF for $h = 2$ is given by the ratio of determinants:

$$
\phi_{22} = \frac{\text{detNum}}{\text{detDen}},
$$

where:

-   The numerator is the determinant of the matrix:

$$
\begin{vmatrix}
1 & \rho_1 \\
\rho_1 & \rho_2
\end{vmatrix},
$$

-   The denominator is the determinant of the matrix:

$$
\begin{vmatrix}
1 & \rho_1 \\
\rho_1 & 1
\end{vmatrix}.
$$

For the numerator:

$$ \text{detNum} = \rho_2 - \rho_1^2. $$

For the denominator, the determinant is:

$$
\text{detDen} = 1 - \rho_1^2.
$$

Substituting the determinants into the formula for $\phi_{22}$:

$$
\phi_{22} = \frac{\rho_2 - \rho_1^2}{1 - \rho_1^2}.
$$

Now, substituting$\rho_1$ and $\rho_2$ from the autocovariance
calculations:

$$
\rho_1 = \frac{\alpha_1 (1 + \alpha_2)}{1 + \alpha_1^2 + \alpha_2^2}, \quad \rho_2 = \frac{\alpha_2}{1 + \alpha_1^2 + \alpha_2^2}.
$$

Thus:

$$
\phi_{22} = \frac{\frac{\alpha_2}{1 + \alpha_1^2 + \alpha_2^2} - \left( \frac{\alpha_1 (1 + \alpha_2)}{1 + \alpha_1^2 + \alpha_2^2} \right)^2}{1 - \left( \frac{\alpha_1 (1 + \alpha_2)}{1 + \alpha_1^2 + \alpha_2^2} \right)^2}.
$$

After substituting the expressions for $\rho_1$ and $\rho_2$, we obtain
the final compact expression for $\phi_{22}$:

$$
\phi_{22} = \frac{\alpha_2 (1 + \alpha_1^2 + \alpha_2^2) - \alpha_1^2 (1 + \alpha_2)^2}{(1 + \alpha_1^2 + \alpha_2^2)^2 - \alpha_1^2 (1 + \alpha_2)^2}.
$$

**D.** The chosen values for $\alpha_1$ and $\alpha_2$ are 0.7 and 0.2,
respectively. These values were selected for two main reasons: to
represent a decay in the influence of past random shocks $S_t$ on $X_t$
and to ensure the stability of the process.

The higher value of $\alpha_1$ (0.7) compared to $\alpha_2$ (0.2)
indicates that the most recent shock $S_{t-1}$ has a greater impact on
the current state of the process than older shocks $S_{t-2}$. In other
words, the most recent information is the most influential.

Although the influence of past shocks decreases over time, it is still
present. This is represented by the value of $\alpha_2$, which, although
smaller than $\alpha_1$, is still different from zero.

The stability of the process is guaranteed by the fact that the sum of
the absolute values of $\alpha_1$ and $\alpha_2$ is less than 1:

$$
|\alpha_1| + |\alpha_2| = 0.7 + 0.2 = 0.9
$$

In summary, the chosen values for $\alpha_1$ and $\alpha_2$ allow for
the creation of a realistic and stable model, where past shocks have a
decreasing influence on the current state of the process.

Then, by the invertibility's property, the coefficients must respect
these conditions:

$$ \begin{cases}     \alpha_2 + \alpha_1 < 1, \\    \alpha_2 - \alpha_1 < 1, \\    -1 < \alpha_2 < 1.\end{cases} $$

**E.**

```{r, echo=TRUE}
# Define the parameters
alpha1 <- 0.7  # Coefficient for S_(t-1)
alpha2 <- 0.2  # Coefficient for S_(t-2)

# Define the length T of the process
T <- 100  # Set the length of the time series

# Define the outcomes S_t (replace with actual data if available)
set.seed(42)  # For reproducibility
S <- sample(0:2, T, replace = TRUE)  # Random outcomes for S_t

# Initialize X_t with zeros
X <- numeric(T)  # Pre-allocate space for X_t
X[1:2] <- S[1:2]  # Use the first two S values directly for X[1] and X[2]

# Generate X_t based on the process
for (t in 3:T) {
  X[t] <- alpha1 * S[t-1] + alpha2 * S[t-2] + S[t]
}

# Plot the generated time series
plot(X, type = "l", col = "purple", lwd = 2, xlab = "Time", ylab = expression(X[t]),
     main = "Generated Time Series for X_t")
grid()
```

**F.**

```{r}
# Compute sample statistics
sample_mean <- mean(X)
sample_variance <- var(X)
sample_acf <- acf(X, plot = FALSE)$acf[1:4]  # lags: 0, 1, 2, 3
```

```{r}
# Print results
cat("Sample Mean:", round(sample_mean, 2), "\n")
```

```{r}
cat("Sample Variance:", round(sample_variance, 2), "\n")
```

```{r}
cat("Sample Autocorrelations:", round(sample_acf, 2), "\n")
```

The MA(2) process is defined as:

$$
X_t = \epsilon_t + \alpha_1 \epsilon_{t-1} + \alpha_2 \epsilon_{t-2}
$$

where: - $X_t$ is the value of the time series at time $t$, -
$\epsilon_t$ is white noise with zero mean and variance
$\sigma^2_\epsilon$, - $\alpha_1 = 0.7$ and $\alpha_2 = 0.2$ are the
parameters of the process.

Considering the results found previously, we can finally compute:

**The mean of** $X_t$ given by:

$$
E[X_t] = \alpha_1 + \alpha_2 + 1
$$

Substituting $\alpha_1 = 0.7$ and $\alpha_2 = 0.2$:

$$
E[X_t] = 0.7 + 0.2 + 1 = 1.9
$$

Thus, the mean is $E[X_t] = 1.9$.

The sample mean ($1.53$) is very close to the new mean of $1.9$,
indicating that the time series is somewhat centered around the expected
value.

**The variance** of $X_t$ is:

$$
\text{Var}(X_t) = \text{Var}(\epsilon_t + \alpha_1 \epsilon_{t-1} + \alpha_2 \epsilon_{t-2})
$$

Since $\epsilon_t$, $\epsilon_{t-1}$, and $\epsilon_{t-2}$ are
uncorrelated:

$$
\text{Var}(X_t) = \text{Var}(\epsilon_t) + \alpha_1^2 \text{Var}(\epsilon_{t-1}) + \alpha_2^2 \text{Var}(\epsilon_{t-2})
$$

Given that the variance of each $\epsilon_t$ is $\sigma^2$, we have:

$$
\text{Var}(X_t) = \sigma^2 + \alpha_1^2 \sigma^2 + \alpha_2^2 \sigma^2
$$

$$
\text{Var}(X_t) = \sigma^2 (1 + \alpha_1^2 + \alpha_2^2)
$$

Substituting the values $\alpha_1 = 0.7$ and $\alpha_2 = 0.2$:

$$
\text{Var}(X_t) = \sigma^2 (1 + 0.7^2 + 0.2^2) = \sigma^2 (1 + 0.49 + 0.04) = \sigma^2 \cdot 1.53
$$

Thus, the theoretical variance is $\text{Var}(X_t) = 1.53 \sigma^2$.
Considering that $\sigma^2$ is 1, the value will be $1.53$

The sample variance ($0.93$) is relatively close to the theoretical
variance of $1.53$, suggesting that the variability of the series
matches the theoretical model.

The **autocorrelations** process are:

#### Lag $h = 1$:

$$ \rho_1 = \frac{\alpha_1(1 + \alpha_2)}{1 + \alpha_1^2 + \alpha_2^2} \approx 0.54. $$

***Lag ( h = 2 )**:*
$$ \rho_2 = \frac{\alpha_2}{1 + \alpha_1^2 + \alpha_2^2} \approx 0.37. $$

***Lag ( h = 3 ):*** $$ \rho_3 = 0. $$

***For ( h \> 3 ):*** $$ \rho_h = 0. $$

**Mean:** The sample mean of $1.53$ is close to the theoretical mean of
$1.9$, indicating that the process is approximately centered around a
stable mean. The difference is small and within the expected range of
sampling variability.

**Variance:** The sample variance of $0.93$ is slightly higher than the
theoretical variance of $0.725$, suggesting a minor increase in
variability. However, the variance remains reasonably consistent over
time, as expected for a stationary process.

**Autocorrelations:** - **Lag** $h = 1$: The sample autocorrelation of
$0.49$ is slightly lower than the theoretical value of $0.54$,
indicating a small underestimation of the dependence between consecutive
observations. - **Lag** $h = 2$: The sample autocorrelation of $-0.09$
deviates from the theoretical value of $0.37$. Although the difference
is more noticeable, it remains within the acceptable range for sampling
variation. - **Lag** $h = 3$: The sample autocorrelation of $-0.19$
significantly deviates from the theoretical value of $0$. Such
deviations are expected at higher lags due to reduced theoretical
influence and greater sampling variability.

**G.** Considering an $\alpha_1=1$, we will have different results and
some properties will be not respected.

In particular, as the conditions mentioned before on the point d, we
will note that the property of the invertibility will be no ensured.

$$ \begin{cases}     \alpha_2 + \alpha_1 < 1, \\    \alpha_2 - \alpha_1 < 1, \\    -1 < \alpha_2 < 1.\end{cases} $$

Indeed, with the new value of $\alpha_1$, the process is no longer
invertible.
